Runtime Feedback in a Meta-Tracing JIT for Efficient Dynamic Languages

### 2011-03-30 talk/icooolps2011/paper.tex
add some benchmarks results, of course now we are over the page limit

It has long been an objective of the partial evaluation community to
automatically produce compilers from interpreters. There has been a recent
renaissance of this idea around the approach of tracing just-in-time
compilers. A number of projects have attempted this approach. SPUR \cite{bebenita_spur:_2010} is
a tracing JIT for .NET together with a JavaScript implementation in C\#. PyPy
\cite{armin_rigo_pypys_2006} contains a tracing JIT for RPython (a restricted
subset of Python). This JIT is then used to trace a number of languages
implementations written in RPython. A number of other experiments in this
directions were done, such as an interpreter for Lua in JavaScript, which is run
on and optimized with a tracing JIT for JavaScript
\cite{yermolovich_optimization_2009}.

@@

>>>>

It has long been an objective of the partial evaluation community to
automatically produce compilers from interpreters. There has been a recent
renaissance of this idea around the approach of tracing just-in-time
compilers. A number of projects have attempted this approach. SPUR \cite{bebenita_spur:_2010} is
a tracing JIT for .NET together with a JavaScript implementation in C\#. PyPy
\cite{armin_rigo_pypys_2006} contains a tracing JIT for Python (a restricted
subset of Python). This JIT is then used to trace a number of languages
implementations written in RPython. A number of other experiments in this
directions were done, such as an interpreter for Lua in JavaScript, which is run
on and optimized with a tracing JIT for JavaScript
\cite{yermolovich_optimization_2009}.

@@

For space reasons we cannot perform a full evaluation here, but still want to
present some benchmark numbers. We chose to present two benchmarks, a port of
the classical Richards benchmark in RPython and a Python version of the Telco
decimal benchmark\footnote{\texttt{http://speleotrove.com/decimal/telco.html}},
using a pure Python decimal floating point implementation.

The benchmarks were run on an otherwise idle Intel Core2 Duo P8400 processor
with 2.26 GHz and 3072 KB of cache on a machine with 3GB RAM running Linux
2.6.35. We compared the performance of various Python implementations on the
benchmarks. As a baseline, we used the standard Python implementation in C,
CPython 2.6.6\footnote{\texttt{http://python.org}}, which uses a bytecode-based
interpreter. We compare it against four versions of PyPy's Python interpreter,
all of them with JIT enabled. The PyPy baseline does not enable maps or type
versions. Then we have a version each where maps and versions are enabled alone
and finally a version with both.

All benchmarks were run 50 times in the same process, to give the JIT time to
produce machine code. The arithmetic mean of the times of the last 30 runs were
used as the result. The errors were computed using a confidence interval with a
95\% confidence level \cite{georges_statistically_2007}. The results are
reported in Figure~\ref{fig:times}.

XXX analysis

\begin{figure}
{\footnotesize
\begin{center}
\begin{tabular}{|l|r|r|}
\hline
        &richards[ms]       &telco[ms] \\
\hline
CPython &357.79 $\pm$ 1.32     &1209.67 $\pm$ 2.20\\
speedup &1.00 $\times$ &1.00 $\times$\\
\hline
JIT baseline    &421.87 $\pm$ 0.48     &738.18 $\pm$ 3.29\\
speedup &0.85 $\times$ &1.64 $\times$\\
\hline
JIT map &382.88 $\pm$ 4.40     &834.19 $\pm$ 4.91\\
speedup &0.93 $\times$ &1.45 $\times$\\
\hline
JIT version     &49.87 $\pm$ 0.29      &157.88 $\pm$ 1.79    \\
speedup &7.17 $\times$ &7.66 $\times$\\
\hline
JIT full        &17.89 $\pm$ 1.15      &153.48 $\pm$ 1.86    \\
speedup &20.00 $\times$        &7.88 $\times$\\
\hline
\end{tabular}
\end{center}
}
\caption{Benchmark Results}
\label{fig:times}
\end{figure}
